\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-09-04 NAR_DRAFT.tex
flailing

\subsection{Real tumor data.}
We utilized two real data sets.
In each, deep sequencing and variant validation were performed with the specific purpose of evaluating tumor variant calling pipelines.
These are among a very small number of tumors which have been sequenced at high depth, and for which variant validation was undertaken.
\citet{Griffith2015} sequenced the whole genome of an acute myeloid leukemia (AML) primary tumor to a depth of $>360\mathrm{X}$, and used targeted sequencing to validate nearly 200,000 mutations.
We downloaded aligned reads for this tumor, and called variants with MuTect 1.1.7 as described in Materials \& Methods.
We estimated a per-base mutation rate for this tumor of $3\cdot10^{-8}$, which is line with the mutation rates generally considered normal in AML.
The area under the precision-recall curve is essentially identical (.995 \& .996) for MuTect and \batcave (Figure \ref{NAR-wes_fig} \textbf{a} and Table \ref{table:01}).

@@

\section{DISCUSSION}
We hope that the availability of this algorithm will have a positive effect on the cost/benefit ratio of deeper sequencing in both research and clinical applications
One size does not fit all with respect to the evidence threshold. Users are interested in true and false positive fractions, and users of different types will have different trade-offs in terms of sensitivity and specificty
Clinical studies often threshold at a vaf, rather than using the evidence threshold of the caller. 272X 5\% cutoff \cite{Parsons2016}, why go so deep?
One problem with this is that vaf and associated evidence is dependent on local depth. This is no different than just having a minimum number of variant reads. Why have a model?
The benefit derived from \batcave varies with both the concentration of the tumor mutation profile and the depth of sequencing.
Deeper sequencing leads to more benefit from the algorithm because there are simply more low frequency, and thus more ambiguous, variants accessible in deeply sequenced tumors.
Deep sequencing is useful \cite{Zehir2017}.
\bkmcomment{the above should be some version of, variant calling is important, deeper is better, we are taking a step in that direction}

@@

>>>>

\subsection{Real tumor data.}
We utilized two real data sets.
In each, deep sequencing and variant validation were performed with the specific purpose of evaluating tumor variant calling pipelines.
These are among a very small number of tumors which have been sequenced at high depth, and for which variant validation was undertaken.
\citet{Griffith2015} sequenced the whole genome of an acute myeloid leukemia (AML) primary tumor to a depth of $>360\mathrm{X}$, and used targeted sequencing to validate nearly 200,000 mutations.
We downloaded aligned reads for this tumor, and called variants with MuTect 1.1.7 as described in Materials \& Methods.
We estimated a per-base mutation rate for this tumor of $3\cdot10^{-8}$, which is line with the mutation rates generally considered normal in AML.
The area under the precision-recall curve is essentially identical (.995 \& .996) for MuTect and \batcave (Figure \ref{NAR-real_data_fig} \textbf{a} and Table \ref{table:01}).

@@

\section{DISCUSSION}

We hope that the availability of this algorithm will have a positive effect on the cost/benefit ratio of deeper sequencing in both research and clinical applications
One size does not fit all with respect to the evidence threshold. Users are interested in true and false positive fractions, and users of different types will have different trade-offs in terms of sensitivity and specificty
Clinical studies often threshold at a vaf, rather than using the evidence threshold of the caller. 272X 5\% cutoff \cite{Parsons2016}, why go so deep?
One problem with this is that vaf and associated evidence is dependent on local depth. This is no different than just having a minimum number of variant reads. Why have a model?
The benefit derived from \batcave varies with both the concentration of the tumor mutation profile and the depth of sequencing.
Deeper sequencing leads to more benefit from the algorithm because there are simply more low frequency, and thus more ambiguous, variants accessible in deeply sequenced tumors.
Deep sequencing is useful \cite{Zehir2017}.
\bkmcomment{the above should be some version of, variant calling is important, deeper is better, we are taking a step in that direction}

@@

\begin{figure}[b]
  \begin{center}
  \includegraphics{figures/aml_plot.pdf}
  \end{center}
  \caption{aml and cell. Need no color and drop calibration}
\label{NAR-real_data_fig}
\end{figure}
